\phatraid, yo

### 2014-05-15 paper/paper.tex
Last updates

\title{PHAT RAID, yo}
\author{Andrew Johnson \and Daniel King \and Lucas Waye \and Scott Moore}
\date{May 13, 2014}

@@

Distributed file stores, like Google's \chubby{} and CS260r's \phat{}, send all
files through a single master node which consistently replicates the file on
many slaves. In this scheme, the master node's throughput is a bottleneck for
storing large files. We present \phatraid{} which mitigates this bottleneck by
partitioning files across many \paxos{} groups which form a \raid{1} group.

@@

Google's \chubby{}\cite{burrows-chubby} is a distributed file system which
achieves consensus in the face of limited failures. Unfortunately, for storing
large files disk latency on the cluster nodes is a bottleneck. We propose
building a RAID group out of distributed file system clusters. Large files are
sharded and disk latency is reduced proportional to number of clusters. Failure
resiliency is not completely sacrificed and we provide an analysis of the
new failure modes.

Our setting is in a data center where we have low network latency, high network
throughput, and the need to store large files with tolerance to some CPU,
network, and disk failure. In this setting, we reason that the major bottleneck
for replicated file storage is disk read/write latency. We combat this
bottleneck with ideas from RAID, which allow for faster disk access, and ideas
from distributed consensus algorithms, which allow for consistent replication.

@@

Figure \ref{fig:hdd-vs-sdd} shows a comparison between read times for a SSD and HDD as a function of position on the disk. HDD storage performance is greatly influenced by where on the disk platter the data was placed. Data can be placed on the outer rim of the platter much more quickly than data placed near the center of the platter as the outer rim is moving faster in comparison (due to conservation of angular momentum). In contrast to HDDs, SSDs do not exhibit this problem as there is no spinning platter. The left diagram in Figure \ref{fig:hdd-vs-sdd} shows relatively constant read performance throughout the entire disk. Additionally, HDDs have difficultly with random accesses as the drive may need to wait for the drive head to be directly over the correct position on the platter. This problem becomes especially apparent when an individual file is fragmented into multiple blocks by the file system. SSDs have relatively better random access times, in general, than HDDs. Sequential access tends to be many times faster than random access for both SSDs and HDDs \cite{disk-perf}. HDD performance can also can suffer from {\em spin-up delay}, which happens as a result of a power-saving measure that stops spinning the disk during idle times.

We decided that HDDs have far too much performance variability in order to be properly modeled. As a result, we instead chose to simulate the performance characteristics of SSDs as they have relatively constant access times regardless of sector position on the drive and also do not suffer from {\em spin-up delay}.

@@

\phatraid{} is implemented in Erlang. Erlang is a pure, functional language with
extensive library support for distributed applications. Erlang has one unit of
concurrency, the \emph{process}, and one unit of distribution, the
\emph{node}. Each Erlang node consists of an arbitrary number of
processes. Erlang nodes may live on the same machine or on different machines
connected by a network. Every Erlang node is given a name by an Erlang name
server. Erlang processes may communicate with any Erlang node or process it can
name. Erlang processes interact with one another by making RPC calls or sending
messages. Identifiers beginning with a lowercase letter are actually atoms, like
Scheme and Lisp's symbols. Identifiers beginning with an uppercase letter are
normal identifiers.

@@

\phatraid{} is implemented in Erlang, reusing our previous implementation of
\phat{}. The implementation is divided into three domains: the server cluster,
the client, and the raid-client. The server code implements a VR-like consensus
algorithm as well as simple file system. The client code exposes a file system
API which hides the communication bookkeeping. The raid-client provides a file
system API which transparently partitions files and reassembles them as they are
sent and received from the server cluster.

@@

Each \phat{} cluster maintains a file system which is separate and distinct from
the file systems of other \phat{} clusters. Generally, two phat clusters do not
communicate. The raid-client and client processes all live on one node. Each
\phat{} master and \phat{} replica lives on its own node.

@@

Instead of sending very large files, we sent smaller strings that were treated by the 
DEAL as being much larger (e.g. 1 character represented 1MB of data). As a result,
we extended the DEAL to perform an additional delay for performing bitwise XOR
on a file as that was the only other operation on a file for our tests. 
In our model, drive delays were much greater than XOR delays (e.g. 20 GB/second 
XOR delay versus 200 MB/second drive delay).

@@

On a client node, the raid-client spawns $C$ client processes which each
maintain a connection to one of the $C$ \phat{} clusters. Each client process
mediates communication between its assigned \phat{} cluster and the raid-client.

@@

\begin{itemize}
\item the supervisor -- restarts the other three processes if any single node dies
\item the server -- passes messages to the VR process if this node is the
  \phat{} master; otherwise, it drops the message and informs the client of the
  current master node
\item VR -- maintains a log of opaque messages, achieving consensus with other
  nodes in the cluster via the Viewstamped Replication protocol
\item the file system -- maintains a simple hierarchical file system with locks
\end{itemize}

@@

The four components are implemented using Erlang behaviors. An Erlang behavior
is a framework which implements common patterns like servers and finite state
machines. In particular, the supervisor uses the \texttt{supervisor} behavior,
the server and file system both use the \texttt{gen\_server} behavior, and VR
uses the \texttt{gen\_fsm} behavior.

\begin{itemize}
\item The \texttt{supervisor} behavior provides a framework for restarting
  failed processes.
\item The \texttt{gen\_server} behavior provides a framework for many-client
  single-server interactions. We implemented functions to process messages
  formatted as Erlang tuples. The \texttt{gen\_server} behavior handles
  low-level socket interaction, messaging queueing, etc.
\item The \texttt{gen\_fsm} behavior generalizes the \texttt{gen\_server}
  behavior by permitting the server to have a finite number of states. Each
  state has a set of functions with which to process messages. This behavior can
  be seen as an ad hoc polymorphic variant of the \texttt{gen\_server} behavior.
\end{itemize}

@@

The \phat{} file system reaches consensus via the Viewstamped Replication
protocol \cite{liskov2012viewstamped}. Listing \ref{lst:genfsm} defines how to
handle prepare messages sent to a node in the replica state. It is piecewise
defined by pattern matching on the arguments. The first two definition clauses
use the \lstinline!when! keyword to specify arbitrary required relationships
between the matched variables.

@@

The call to \lstinline!processBuffer! moves messages to the log, if all previous
messages have now been received. The call to \lstinline!processLog! commits
logged messages which have been newly committed by the master node. Lines
405--408 destructure \lstinline!AfterLog! and bind variables for later use. The
final \lstinline!if! statement responds to the master unless the master has
committed messages we haven't yet received.\footnote{This can happen if the
  network drops a message whose operation number lies between the replica's last
  commit number and the master's last commit number.}

@@

We distinguish between two types of failure: cessation of progress and loss of
data. In a VR cluster without persistent stores, data loss of fully propagated,
commited data only occurs when all $2f+1$ nodes in the cluster fail \footnote{If
  more than $f+1$ nodes die some ``commited but not fully propogated data'' can
  be lost}. In a VR cluster, progress only ceases if $f+1$ nodes are lost. In
\phatraid{} the conditions for cessation of progress and loss of data are
complicated by the combination of RAID and VR.

@@

  \caption{Some Failed Node Distributions of \phatraidcf{3}{1}}
  \label{fig:failures}
\end{figure}

Four arrangements of failed nodes in \phatraidcf{3}{1} are depicted in Figure
\ref{fig:failures}. In this section, ``data is lost'' means that, if no
persistent, fail-safe store is used, some data will be lost. Adding a
persistent, fail-safe store obviously prevents loss of committed
data. Additionally, we assume node failures are independent and uniformly
distributed throughout the \phatraid{} group.

The top left corner, Figure \ref{fig:failures_allok}, depicts five of the nine
nodes failing. This state is \emph{not} a failure state. The group can still
make progress and no data is lost. The two surviving clusters can reach
consensus on new values and can reconstruct the failed cluster's data.

The top right corner, Figure \ref{fig:failures_noprogress}, depicts four of the
nine nodes failing. This state is a failure state. The group cannot make
progress; however, no data can been lost\footnote{Here and later in this paper
  we mean that no data which has been decided and propagated throughout the
  cluster can be lost. In VR, by contrast, if data is decided and propagated
  throughout the cluster, that data can only be lost if every node is lost. We
  also assume the nodes do not have persistent, fail-safe stores.}. Progress is
blocked because more than one cluster cannot reach consensus. Since no cluster
was completely destroyed, all data still exists on at least one node.

The bottom left corner, Figure \ref{fig:failures_impossible}, is impossible
because the group certainly cannot make progress if data has been lost.

The bottom right corner, Figure \ref{fig:failures_noprogress_dataloss}, depicts
six of the nine nodes failing. This state is also a failure state. The group
cannot make progress and will lose data. Progress is blocked because more than
one cluster cannot reach consensus. Additionally, two complete clusters have
been lost so half of the data cannot be recovered.

@@

By the pidgeon hole principle, the upper bound of failed nodes which still
permit progress is:

@@

because $C-1$ clusters can make progress, i.e. at least $f+1$ nodes are up in
each cluster, and only one cluster is completely dead.

@@

because $C-1$ clusters have at least one node up and one cluster has no nodes
up.

@@

In a \phatraidcf{3}{1} group, less than one-quarter of four-failure arrangements
stop progress. The necessary condition for progress cessation is that two or
more clusters have lost at least $f+1$ nodes. By the pidgeon hole principle,
only the two-in-two, four-failure arrangement stops progress. We can count the
number of such arrangements:

@@

When five nodes fail, they can be distributed into each group in three different
ways: (3,1,1), (3,2,0) and (2,2,1). The first case does not stop progress
because only one cluster fails. The later two cases stop progress because two
clusters fail. We calculate the probability of \phatraid{} group failure when
five nodes fail:

@@

The general formula for the failure probability distribution is not clear, but
it is clear that a \phatraidcf{$C$}{$f$} group is absolutely resilient to $2(f +
1) - 1$ or fewer failures. A \phatraidcf{$C$}{$f$} group is likely resilient
to $2(f+1)$ failures and unlikely resistant to $(C-1)\cdot f + 2f + 1$ failures.

@@

For \phatraidcf{6}{1} we calculate:

@@

\begin{figure}
  \includegraphics[scale=0.45]{Store.pdf}
  \caption{Scatter Plot of Various \phatraidcf{$C$}{$f$} Configurations}
  \label{fig:plot}
\end{figure}

We evaluated a prototype \phatraidcf{$C$}{$f$} implementation. We simulated disk
latency with sleeps. All experiments are run on a 2013-model MacBook Pro with
8GB of RAM, and a 3GHz Intel Core i7. We found that latency scales as
$\frac{1}{C-1}$. This is the expected behavior for breaking a file into $C-1$
chunks. We always use one bit of parity.

@@

We've presented RAID on a distributed file system which shows that sharding
files improves latency for large files while maintaining reasonable failure
probabilities.

>>>>

\title{\phatraid, yo}
\author{Andrew Johnson \and Daniel King \and Lucas Waye \and Scott Moore}
\date{May 13, 2014}

@@

Distributed file stores, like Google's \chubby{}\cite{burrows-chubby}
and CS260r's \phat{}\cite{class}, send all files through a single
master node which consistently replicates the file on many slaves. In
this scheme, the master node's throughput is a bottleneck for storing
large files. We present \phatraid{} which mitigates this bottleneck by
partitioning files across many \paxos{} groups which form a \raid{}\cite{raid}
group.

@@

Google's \chubby{}\cite{burrows-chubby} is a distributed file system which
achieves consensus in the face of limited failures. Unfortunately, for storing
large files disk latency on the cluster nodes is a bottleneck. We propose
building a RAID (Redundant Array of Inexpensive Disks)\cite{raid} group out of distributed file system clusters. Large files are
sharded and disk latency is reduced proportional to number of clusters. Failure
resiliency is not completely sacrificed and we provide an analysis of the
new failure modes.

Our setting is in a data center where we have low network latency, high network
throughput, and the need to dependably store large files with tolerance to some CPU,
network, and disk failure. In this setting, we hypothesize that the major bottleneck
for replicated file storage is disk read/write latency. We combat this
bottleneck with ideas from RAID, which allow for faster file access, and ideas
from distributed consensus algorithms, which allow for consistent replication.

@@

Figure \ref{fig:hdd-vs-sdd} shows a comparison between read times for
a SSD and HDD as a function of position on the disk. HDD storage
performance is greatly influenced by where on the disk platter the
data was placed. Data can be placed on the outer rim of the platter
much more quickly than data placed near the center of the platter as
the outer rim is moving faster in comparison (due to conservation of
angular momentum). In contrast to HDDs, SSDs do not exhibit this
problem as there is no spinning platter. The left diagram in Figure
\ref{fig:hdd-vs-sdd} shows relatively constant read performance
throughout the entire disk. Additionally, HDDs have difficultly with
random accesses as the drive may need to wait for the drive head to be
directly over the correct position on the platter. This problem
becomes especially apparent when an individual file is fragmented into
multiple blocks by the file system. SSDs have better random access
times, in general, than HDDs. Sequential access tends to be many times
faster than random access for both SSDs and HDDs \cite{disk-perf}. HDD
performance can also can suffer from {\em spin-up delay}, which
happens as a result of a power-saving measure that stops spinning the
disk during idle times.

We decided that HDDs have far too much performance variability in
order to be properly modeled. As a result, we instead chose to
simulate the performance characteristics of SSDs as they have
relatively constant access times regardless of sector position on the
drive and also do not suffer from {\em spin-up delay}. Also, because
HDDs are slower, and our hypothesis supposes that disk is the
bottleneck, this should be a more challenging case for \phatraid{}.

@@

\phatraid{} is implemented in Erlang. Erlang is a pure, functional language with
extensive library support for distributed applications. Erlang has one unit of
concurrency, the \emph{process}, and one unit of distribution, the
\emph{node}. Each Erlang node consists of an arbitrary number of
processes. Erlang nodes may live on the same machine or on different machines
connected by a network. Every Erlang node is given a name by an Erlang name
server. Erlang processes may communicate with any Erlang node or process it can
name. Erlang processes interact with one another by making RPC calls or sending
messages. Identifiers beginning with a lowercase letter are actually
\emph{atoms}, a literal constant similar to
Scheme and Lisp's symbols. Identifiers beginning with an uppercase letter are
normal identifiers.

@@

\phatraid{} is implemented in Erlang, reusing our previous implementation of
\phat{}. The implementation is divided into three domains: the server cluster,
the client, and the raid-client. The server code implements a variant
of the Viewstamped Replication (VR)~\cite{liskov2012viewstamped} consensus
algorithm as well a file system. The client code exposes a file system
API which hides the communication bookkeeping. The raid-client provides a file
system API which transparently partitions files and reassembles them as they are
sent and received from the server cluster.

@@

The master and each replica in a \phat{} cluster maintains a separate
file system. Two different phat clusters do not communicate
directly. The raid-client and client processes all live on one Erlang
node. Each \phat{} master and \phat{} replica lives on its own Erlang
node along with its file system.

@@

Instead of sending very large files, we sent smaller strings that were treated by the 
DEAL as being much larger (e.g. 1 character represented 1MB of data). As a result,
we extended the DEAL to perform an additional delay for performing bitwise XOR
on a file as that was the only other operation on a file for our tests. 
In our model, drive delays were much greater than XOR delays (e.g. 20 GB/second 
XOR throughput versus 200 MB/second drive throughput).

@@

On a raid-client node, $C$ client processes are spawned each of which
maintains a connection to one of the $C$ \phat{} clusters. Each client process
mediates communication between its assigned \phat{} cluster and the raid-client.

@@

\begin{itemize}
\item the supervisor -- restarts the other three processes if any dies
\item the server -- passes messages to the VR process if this node is the
  \phat{} master; otherwise, it drops the message and informs the client of the
  current master node
\item VR -- maintains a log of opaque messages, achieving consensus with other
  nodes in the cluster via the Viewstamped Replication protocol
\item the file system -- maintains a hierarchical file system with locks
\end{itemize}

@@

The four components are implemented using Erlang behaviors. An Erlang behavior
is a framework which implements common patterns like servers and finite-state
machines. In particular, the supervisor uses the \texttt{supervisor} behavior,
the server and file system both use the \texttt{gen\_server} behavior, and VR
uses the \texttt{gen\_fsm} behavior.

\begin{itemize}
\item The \texttt{supervisor} behavior provides a framework for restarting
  failed processes.
\item The \texttt{gen\_server} behavior provides a framework for many-client
  single-server interactions. We implemented functions to process messages
  formatted as Erlang tuples. The \texttt{gen\_server} behavior handles
  low-level socket interaction, messaging queueing, etc.
\item The \texttt{gen\_fsm} (finite-state machine) behavior
generalizes the \texttt{gen\_server} behavior by permitting the server
to have a finite number of states. Each state has a set of functions
with which to process messages. This behavior can be seen as an ad hoc
polymorphic variant of the \texttt{gen\_server} behavior.
\end{itemize}

@@

The \phat{} file system reaches consensus via the Viewstamped Replication
protocol \cite{liskov2012viewstamped}. Listing \ref{lst:genfsm} defines how to
handle prepare messages sent to a node in the replica state. It is piecewise
defined by pattern matching on the arguments. The first two definition clauses
use the \lstinline!when! keyword to specify arbitrary required relationships
between the matched variables.\footnote{\lstinline{?debugFmt} is a macro
    for printing debug messages when a flag is set}

@@

The call to \lstinline!processBuffer! moves messages to the log, if all previous
messages have now been received. The call to \lstinline!processLog! commits
logged messages which have been newly committed by the master node. Lines
405--408 destructure \lstinline!AfterLog! and bind variables for later use. The
final \lstinline!if! statement responds to the master unless the master has
committed messages the replica has not yet received.\footnote{This can happen if the
  network drops a message whose operation number lies between the replica's last
  commit number and the master's last commit number.}

@@

We distinguish between two types of failure: cessation of progress and loss of
data. In a VR cluster data loss of fully propagated,
commited data only occurs when all $2f+1$ nodes in the cluster fail \footnote{If
  more than $f+1$ nodes die some ``commited but not fully propogated data'' can
  be lost}. In a VR cluster, progress only ceases if $f+1$ nodes are lost. In
\phatraid{} the conditions for cessation of progress and loss of data are
complicated by the combination of RAID and VR.

@@

  \caption{Some Failed Node Distributions of \phatraidcf{3}{1}. Each
    circle represent a singe VR master or replica and each column is a
  single \phat{} cluster. The red lines indicate failures.}
  \label{fig:failures}
\end{figure}

Four arrangements of node failures in \phatraidcf{3}{1} are depicted in
Figure \ref{fig:failures}. There are three columns, each representing
a \phat cluster of three nodes. The red lines indicate failures. We
assume failures are independent and uniformly distributed
throughout the \phatraid{} group.

The top left corner, Figure \ref{fig:failures_allok}, depicts five of the nine
nodes failing. This state is \emph{not} a failure state. The group can still
make progress and no data is lost. The two surviving clusters can reach
consensus on new values and can reconstruct the failed cluster's data
(by taking the XOR of the data on the two surviving clusters).

The top right corner, Figure \ref{fig:failures_noprogress}, depicts four of the
nine nodes failing. This state is a failure state. The group cannot make
progress; however, no data can been lost.\footnote{Here and later in this paper
  we mean that no data which has been decided and propagated throughout the
  cluster can be lost. In VR and normal \phat, by contrast, if data is decided and propagated
  throughout the cluster, that data can only be lost if every node is lost.} Progress is
blocked because more than one cluster is unable to reach consensus. Since no cluster
was completely destroyed, all data still exists on at least one node.

The bottom left corner, Figure \ref{fig:failures_impossible}, is impossible
because the group cannot make progress if data has been lost.

The bottom right corner, Figure \ref{fig:failures_noprogress_dataloss}, depicts
six of the nine nodes failing. This state is also a failure state. The group
cannot make progress and will lose data. Progress is blocked because more than
one cluster is unable to reach consensus. Additionally, two complete clusters have
been lost so at least half of the data cannot be recovered.\footnote{If
  only the cluster with the parity bit remains then all useable data
  has been lost}

@@

The upper bound of failed nodes which still
permit progress is:

@@

because $C-1$ clusters could still make progress, e.g. if at least $f+1$ nodes are up in
each cluster, and only one cluster is completely dead.

@@

because it is possible that $C-1$ clusters have at least one node up
and only one cluster has no nodes up.

@@

In a \phatraidcf{3}{1} group, less than one-quarter of four-failure arrangements
stop progress. The necessary condition for progress cessation is that two or
more clusters have lost at least $f+1$ nodes.
Only the two-in-two, four-failure arrangement stops progress. We can count the
number of such arrangements:

@@

When five nodes fail, they can be distributed into each cluster in three different
ways: (3,1,1), (3,2,0) and (2,2,1). The first case does not stop progress
because only one cluster fails. The later two cases stop progress because two
clusters fail. We calculate the probability of \phatraid{} group failure when
five nodes fail:

@@

The general formula for the failure probability distribution is left
as an exercise for the reader, but
it is clear that a \phatraidcf{$C$}{$f$} group is absolutely resilient to $2(f +
1) - 1$ or fewer failures. In addition, a \phatraidcf{$C$}{$f$} group is likely resilient
to $2(f+1)$ failures and unlikely resistant to $(C-1)\cdot f + 2f + 1$ failures.

@@

For example, for a \phatraidcf{6}{1} we calculate:

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{Store.png}
  \caption{Scatter Plot of Various \phatraidcf{$C$}{$f$} Configurations}
  \label{fig:plot}
\end{figure}

We evaluated a prototype \phatraidcf{$C$}{$f$} implementation. We simulated disk
latency with sleeps. All experiments are run on a 2013-model MacBook Pro with
8GB of RAM, and a 3GHz Intel Core i7. We found that latency scales asp
$\frac{1}{C-1}$. This is the expected behavior for breaking a file into $C-1$
chunks. We always use one bit of party.

@@

We've presented RAID on a distributed file system which shows that sharding
files can improve latency for large files while maintaining reasonable failure
probabilities.
